---
title: "Report2"
format: html
editor: visual
---

#General

```{r, general setup}
library('nlme')
library('lme4')
library('ggplot2')
library('clubSandwich')

ebas <- "sum_entropy_child~study+sum_entropy_parent+sum_entropy_parent+ntokens_child+age_months_t2+sex_child+age_months_t2:sex_child"
sbas <- "sum_surprisal_child~study+sum_surprisal_parent+ntokens_child+age_months_t2+sex_child+age_months_t2:sex_child"
mbas <- "ntokens_child~study+ntokens_parent+age_months_t2+sex_child+age_months_t2:sex_child"

edif <- "+age_diff+sum_entropy_parent:age_diff"
sdif <- "+age_diff+sum_surprisal_parent:age_diff"
mdif <- "+age_diff+ntokens_parent:age_diff"

etok <- "+ntokens_parent+sum_entropy_parent:ntokens_parent"
stok <- "+ntokens_parent+sum_surprisal_parent:ntokens_parent"

etd <- "+ntokens_parent:age_diff+sum_entropy_parent:ntokens_parent:age_diff"
std <- "+ntokens_parent:age_diff+sum_surprisal_parent:ntokens_parent:age_diff"

bs <- "+ (1|name)"

```

```{r, function general setup}
pasteq <- function(eqparts){
  peq <- paste(eqparts,collapse = "") 
  return(as.formula(peq))
}

qqplot_with_hist <- function(model, xlab, ylab, hist_frac = 1) {
  res <- resid(model, type = "normalized")
  qq  <- qqnorm(res, plot.it = FALSE)  # qq$x = theoretical, qq$y = sample (residuals)

  # note: res on x-axis now, theoretical quantiles on y-axis
  plot(qq$y, qq$x, type = "n", xlab = xlab, ylab = ylab,
       bty = "n", main = "")

  usr <- par("usr")  # c(xmin, xmax, ymin, ymax) of current plot

  # histogram of the residuals (these become vertical bars along the bottom)
  h <- hist(res, breaks=20, plot = FALSE)

  # scale bar heights so the tallest bar takes up `hist_frac` of the y-range,
  # anchored to the bottom edge of the plot
  bar_len <- h$counts / max(h$counts) * diff(usr[3:4]) * hist_frac

  rect(xleft   = h$breaks[-length(h$breaks)],
       xright  = h$breaks[-1],
       ybottom = usr[3],
       ytop    = usr[3] + bar_len,
       col = adjustcolor("grey70", alpha.f = 0.5),
       border = NA)

  # reference line and QQ points on top
  abline(0, 1, col = "red")
  points(qq$y, qq$x, pch = 1)
}

runmodel <- function(tempeq,data,show_vif = TRUE){
  model <-lme(pasteq(tempeq),random = ~ 1  | name, 
            data = data, na.action = na.omit)
  #model0 <- gls(pasteq(tempeq), data = data, na.action = na.omit)
  #print(anova(model0,model))
  #tmodel <- round(summary(model)$tTable,3)
  #print(tmodel[!grepl("^study", rownames(tmodel)), ])
  
  V_cr <- vcovCR(model, type = "CR2")   # CR2 recommended (bias-reduced, best small-sample behavior)
  ctest <- coef_test(model, vcov = V_cr, test = "Satterthwaite")
  rownames(ctest) <- ctest$Coef
  ctest <- round(as.matrix(ctest[ , -1]), 3)
  print(ctest[!grepl("^study", rownames(ctest)), ])
  
  print(plot(Variogram(model, form = ~age_months_t2),
       xlab = "Distance (mois)", 
       ylab = "Semi-variogramme",
       main = ""))
  #acf_fit <- ACF(model, resType = "normalized")
  #print(plot(acf_fit, alpha = 0.05))  # adds confidence bands
  print(plot(model, xlab = "Valeurs ajustées", 
       ylab = "Résidus standardisés",
       main = ""))
  #hist(resid(model))
  qqplot_with_hist(model,
                  xlab = "Résidus standardisés et quantiles",
                  ylab = "Quantiles d'une courbe normale standard")
  
  if(show_vif){
    tvif <- lmer(pasteq(c(tempeq,bs)),data = data, na.action = na.omit)
    car::vif(tvif)
  }
}

```

```{r, data setup}
nc_data <- read.csv('/home/hereinlies/Documents/Documents/Ecole/HEC/ProjetSupervise/data/Configs/NoContext1/new_model.csv')
nc_data$study <- factor(nc_data$study)
nc_data$sex_child <- factor(nc_data$sex_child, levels = c('male','female'))

c3_data <- read.csv('/home/hereinlies/Documents/Documents/Ecole/HEC/ProjetSupervise/data/Configs/Context3_1/new_model.csv')
c3_data$study <- factor(c3_data$study)
c3_data$sex_child <- factor(c3_data$sex_child, levels = c('male','female'))

c9_data <- read.csv('/home/hereinlies/Documents/Documents/Ecole/HEC/ProjetSupervise/data/Configs/Context9_1/new_model.csv')
c9_data$study <- factor(c9_data$study)
c9_data$sex_child <- factor(c9_data$sex_child, levels = c('male','female'))

c27_data <- read.csv('/home/hereinlies/Documents/Documents/Ecole/HEC/ProjetSupervise/data/Configs/Context27_1/new_model.csv')
c27_data$study <- factor(c27_data$study)
c27_data$sex_child <- factor(c27_data$sex_child, levels = c('male','female'))
```

```{r, heatmap, fig.width=4.16, fig.height=3.2}
#{r, fig.width=8, fig.height=6, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Histograms and boxplots for key variables", fig.pos='H'}

make_heatmap <- function(data){

  my_colors <- colorRampPalette(c("cyan","darkblue","white","darkred", "yellow"))(20)
  
  cor_mat <- cor(data[, c(3,4,5,6,13,14,15,21)])
  
  # Cluster based on dissimilarity (1 - correlation is common, or use dist() on the matrix)
  hc <- hclust(as.dist(1 - cor_mat))
  ord <- hc$labels[hc$order]
  
  # Convert to long format (base R, no melt)
  cor_df <- as.data.frame(as.table(cor_mat))
  names(cor_df) <- c("Var1", "Var2", "value")
  
  # Reorder factor levels according to clustering
  cor_df$Var1 <- factor(cor_df$Var1, levels = ord)
  cor_df$Var2 <- factor(cor_df$Var2, levels = ord)
  
  label_map <- c(
    "ntokens_parent" = expression(LME[p]),
    "sum_entropy_parent" = expression("Somme entropie"[p]),
    "sum_surprisal_parent" = expression("Somme surprise"[p]),
    "age_diff" = "Temps",
    "sum_entropy_child" = expression("Somme entropie"[e]),
    "sum_surprisal_child" = expression("Somme surprise"[e]),
    "ntokens_child" = expression(LME[e]),
    "age_months_t2" = expression(paste(Age["t2"],"(mois)"))
  )
  
  ggplot(cor_df, aes(x = Var1, y = Var2, fill = value)) +
    geom_tile() +
    scale_fill_gradientn(colors = my_colors, limits = c(-1, 1), name = expression(rho)) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = "", y = "") +
    scale_x_discrete(labels = label_map) +
    scale_y_discrete(labels = label_map)
}

make_heatmap(nc_data)
make_heatmap(c3_data)
make_heatmap(c9_data)
make_heatmap(c27_data)

```

#Report test 
##Entropy

```{r, entropy model}
runmodel(c(ebas),nc_data,show_vif = FALSE)       #simplest
runmodel(c(ebas),c3_data,show_vif = FALSE)       #simplest
runmodel(c(ebas),c9_data,show_vif = FALSE)       #simplest
runmodel(c(ebas),c27_data,show_vif = FALSE)       #simplest
```

##Surprisal

```{r, entropy model}
runmodel(c(sbas),nc_data,show_vif = FALSE)       #simplest
runmodel(c(sbas),c3_data,show_vif = FALSE)       #simplest
runmodel(c(sbas),c9_data,show_vif = FALSE)       #simplest
runmodel(c(sbas),c27_data,show_vif = FALSE)       #simplest
```

##MLU

```{r, entropy model}
runmodel(c(mbas),nc_data,show_vif = FALSE)       #simplest
runmodel(c(mbas),c3_data,show_vif = FALSE)       #simplest
runmodel(c(mbas),c9_data,show_vif = FALSE)       #simplest
runmodel(c(mbas),c27_data,show_vif = FALSE)       #simplest
```

#Exploratory 
## No context

### Entropy

```{r, no context entropy model}
runmodel(c(ebas,edif,etok,etd),nc_data) #full
runmodel(c(ebas,edif),nc_data)  #with dif
runmodel(c(ebas,etok),nc_data)  #with tok
runmodel(c(ebas),nc_data)       #simplest
```

### surprisal

```{r, no context suprisal model}
runmodel(c(sbas,sdif,stok,std),nc_data) #full
runmodel(c(sbas,sdif),nc_data)  #with dif
runmodel(c(sbas,stok),nc_data)  #with tok
runmodel(c(sbas),nc_data)       #simplest
```

### MLU

```{r, no context mlu model}
runmodel(c(mbas,mdif),nc_data)  #with dif
runmodel(c(mbas),nc_data)       #simplest
```

### Extra code

modelmlu1 \<-lme(ntokens_child\~study+ntokens_parent +age_months_t2+age_diff+sex_child+age_months_t2:sex_child+ntokens_parent:age_diff +ntokens_parent:age_diff,\
random = \~ 1 \| name, data = nc_data, na.action = na.omit, correlation = corLin(form = \~age_months_t2 \| name))

modelmlu2 \<-lme(ntokens_child\~study+ntokens_parent +age_months_t2+age_diff+sex_child+age_months_t2:sex_child+ntokens_parent:age_diff +ntokens_parent:age_diff,\
random = \~ 1 \| name, data = nc_data, na.action = na.omit)

modelmlu3 \<-lme(ntokens_child\~study+ntokens_parent +age_months_t2+age_diff+sex_child+age_months_t2:sex_child+ntokens_parent:age_diff +ntokens_parent:age_diff,\
random = \~ age_months_t2 \| name, data = nc_data, na.action = na.omit, correlation = corLin(form = \~age_months_t2 \| name))

plot(Variogram(modelmlu2, form = \~ age_months_t2 \| name, resType = "pearson"))\
plot(Variogram(modelmlu3, form = \~ age_months_t2 \| name, resType = "normalized"))

## Context 3

### Entropy

```{r, no context entropy model}
runmodel(c(ebas,edif,etok,etd),c3_data) #full
runmodel(c(ebas,edif),c3_data)  #with dif
runmodel(c(ebas,etok),c3_data)  #with tok
runmodel(c(ebas),c3_data)       #simplest
```

### surprisal

```{r, no context suprisal model}
runmodel(c(sbas,sdif,stok,std),c3_data) #full
runmodel(c(sbas,sdif),c3_data)  #with dif
runmodel(c(sbas,stok),c3_data)  #with tok
runmodel(c(sbas),c3_data)       #simplest
```

### MLU

```{r, no context mlu model}
runmodel(c(mbas,mdif),c3_data)  #with dif
runmodel(c(mbas),c3_data)       #simplest
```

## Context 9

### Entropy

```{r, no context entropy model}
runmodel(c(ebas,edif,etok,etd),c9_data) #full
runmodel(c(ebas,edif),c9_data)  #with dif
runmodel(c(ebas,etok),c9_data)  #with tok
runmodel(c(ebas),c9_data)       #simplest
```

### surprisal

```{r, no context suprisal model}
runmodel(c(sbas,sdif,stok,std),c9_data) #full
runmodel(c(sbas,sdif),c9_data)  #with dif
runmodel(c(sbas,stok),c9_data)  #with tok
runmodel(c(sbas),c9_data)       #simplest
```

### MLU

```{r, no context mlu model}
runmodel(c(mbas,mdif),c9_data)  #with dif
runmodel(c(mbas),c9_data)       #simplest
```

## Context 27

### Entropy

```{r, no context entropy model}
runmodel(c(ebas,edif,etok,etd),c27_data) #full
runmodel(c(ebas,edif),c27_data)  #with dif
runmodel(c(ebas,etok),c27_data)  #with tok
runmodel(c(ebas),c27_data)       #simplest
```

### surprisal

```{r, no context suprisal model}
runmodel(c(sbas,sdif,stok,std),c27_data) #full
runmodel(c(sbas,sdif),c27_data)  #with dif
runmodel(c(sbas,stok),c27_data)  #with tok
runmodel(c(sbas),c27_data)       #simplest
```

### MLU

```{r, no context mlu model}
runmodel(c(mbas,mdif),c27_data)  #with dif
runmodel(c(mbas),c27_data)       #simplest
```
